Parallel Multipoint Approximation Method for Large Scale Optimization Problems \thanks{This study was supported by the Russian Science Foundation, project No. 16-11-10150.}

### 2017-12-01 2018/PAVT_Toropov/PAVT2018_MAM.tex
minor corrections

\author{ Victor P. Gergel%(ORCID 0000-0002-4013-2329)
$^1$ \and
Konstantin A. Barkalov%(ORCID 0000-0001-5273-2471)
$^1$ \and
Evgeny A. Kozinov%(ORCID 0000-0001-6776-0096)
$^1$ \and
Vassili V. Toropov%(ORCID 0000-0002-7017-5055)
$^{1,2}$
}
%
\authorrunning{V.P. Gergel \and K.A. Barkalov \and E.A.Kozinov \and V.V. Toropov}
% (feature abused for this document to repeat the title also on left hand pages)

@@

Second, one can correct the algorithm with the purpose of parallel computing several values of the objective function and constraints at different points of the search domain. According to the MAM rules, in design of experiments $P$ 
sampling points are formed in current trust region at every iteration. The function values at these points can be computed on different processors (nodes) in parallel. The above corresponds to the parallelization of Step 2a of the algorithm with the use of the distributed memory. The number of sampling points generated within each iteration may be set equal to $P = k\cdot NP$ number, where $NP$ is a number of available processors (or nodes) and $k \geq 1$. In  terms of the time, the latter will be equivalent to $NP$ function evaluations per step. This method has been implemented successfully in \cite{PolynkinToropov2008} and has demonstrated a good efficiency since at the number of the design variables of the order of $10^2$ major time is consumed by the function evaluations, the work of the MAM itself (in the sequential regime) introduces minor overhead.

However, when the number of variables becomes of the order of $10^3$, the work of the sequential part of MAM begins to affect the total problem solving time essentially (assuming the time of computing the objective function 
values remains constant). Thus, for the problem considered in Section 4 the time of execution of a single iteration of the method increased more than 4000 times (from 1.5 up to 6000 seconds) when increasing the number of variables from 100 up to  1000. Therefore, another approach to the parallelization of the algorithm has been applied within the framework of the 
present study: namely, the computational rules of MAM  providing the construction of the approximating, the solving of the approximated problem, and the choice of the next trust region (Steps 2b, 2c, and 2d of the algorithm) have been parallelized.

>>>>

\author{ Victor P. Gergel%(ORCID 0000-0002-4013-2329)
$^1$ \and
Konstantin A. Barkalov%(ORCID 0000-0001-5273-2471)
$^1$ \and
Evgeny A. Kozinov%(ORCID 0000-0001-6776-0096)
$^1$ \and \\
Vassili V. Toropov%(ORCID 0000-0002-7017-5055)
$^{1,2}$
}
%
\authorrunning{V.P. Gergel \and K.A. Barkalov \and E.A.Kozinov \and V.V. Toropov}
% (feature abused for this document to repeat the title also on left hand pages)

@@

Second, one can correct the algorithm with the purpose of parallel computing several values of the objective function and constraints at different points of the search domain. According to the MAM rules, in design of experiments $P$ 
sampling points are formed in current trust region at every iteration. The function values at these points can be computed on different processors (nodes) in parallel. The above corresponds to the parallelization of Step 2a of the algorithm with the use of the distributed memory. The number of sampling points generated within each iteration may be set equal to $P = k\cdot NP$ number, where $NP$ is a number of available processors (or nodes) and $k \geq 1$. In  terms of the time, the latter will be equivalent to $NP$ function evaluations per step. This method has been implemented successfully in \cite{PolynkinToropov2008} and has demonstrated a good efficiency since at the number of the design variables of the order of 100 major time is consumed by the function evaluations, the work of the MAM itself (in the sequential regime) introduces minor overhead.

However, when the number of variables becomes of the order of 1000, the work of the sequential part of MAM begins to affect the total problem solving time essentially (assuming the time of computing the objective function 
values remains constant). Thus, for the problem considered in Section 4 the time of execution of a single iteration of the method increased more than 4000 times (from 1.5 up to 6000 seconds) when increasing the number of variables from 100 up to  1000. Therefore, another approach to the parallelization of the algorithm has been applied within the framework of the 
present study: namely, the computational rules of MAM  providing the construction of the approximating, the solving of the approximated problem, and the choice of the next trust region (Steps 2b, 2c, and 2d of the algorithm) have been parallelized.
